---
title: "Tugas Minggu 4 MPDW"
author: "Nabil Bintang Prayoga"
date: "2024-09-13"
output: rmdformats::readthedown
---

# White Noise

```{r}
set.seed(2024)
wn <- rnorm(300)
ts.plot(wn)
```

# MA(2)

```{r}
set.seed(1234)
ma <- wn[c(1,2)]
```

```{r}
for(i in 3:300){
   ma[i] <- wn[i] + 0.4 * wn[i - 1] + 0.6 * wn[i - 2]
}
ma
```

```{r}
ma2 <- arima.sim(list(order=c(0,0,2), ma=c(0.4, 0.6)), n=300)
ma2
```

## Karakteristik MA(2)

### Plot Time Series

```{r}
ts.plot(ma)
```

### Plot ACF

```{r}
acf(ma,lag.max = 20)
```

### Plot PACF

```{r}
pacf(ma)
```

### Plot EACF

```{r}
TSA::eacf(ma)
```

### Scatterplot Antar Lag

#### Korelasi antara $Y_t$ dengan $Y_{t-1}$

```{r}
yt_ma <- ma[-1]
yt_ma

yt_1_ma <- ma[-300]
yt_1_ma
```

```{r}
plot(y=yt_ma,x=yt_1_ma)
```

```{r}
cor(yt_ma,yt_1_ma)
```

#### Korelasi antara $Y_t$ dengan $Y_{t-2}$

```{r}
yt_ma2 <- ma[-c(1,2)]
yt_ma2

yt_2_ma <- ma[-c(299,300)]
yt_2_ma
```

```{r}
plot(y=yt_ma2,x=yt_2_ma)
```

```{r}
cor(yt_ma2,yt_2_ma)
```


#### Korelasi antara $Y_t$ dengan $Y_{t-3}$

```{r}
yt_ma3 <- ma[-c(1,2,3)]
yt_ma3

yt_3_ma <- ma[-c(288, 299,300)]
yt_3_ma
```

```{r}
plot(y=yt_ma3,x=yt_3_ma)
```

```{r}
cor(yt_ma3,yt_3_ma)
```

# AR(2)

```{r}
set.seed(1234)
```

```{r}
n<-length(wn)
n
ar <- c(1:n) 
for (i in 3:n) {ar[i]<-wn[i]+0.5*ar[i-1]+0.2*ar[i-2]}
ar
```

```{r}
ar1 <- arima.sim(list(order=c(2,0,0), ar=c(0.5,0.2)), n=300)
ar1
```

## Karakteristik AR(2)

### Plot Time Series

```{r}
ts.plot(ar)
```

### Plot ACF

```{r}
acf(ar)
```

### Plot PACF

```{r}
pacf(ar)
```

### Plot EACF (Extended)

```{r}
TSA::eacf(ar)
```

### Scatterplot Antar Lag

#### Korelasi antara $Y_t$ dengan $Y_{t-1}$

```{r}
yt_ar <- ar[-1]
yt_ar

yt_1_ar <- ar[-300]
yt_1_ar
```

```{r}
plot(y=yt_ar,x=yt_1_ar)
```

```{r}
cor(yt_ar,yt_1_ar)
```

#### Korelasi antara $Y_t$ dengan $Y_{t-2}$

```{r}
yt_ar2 <- ar[-c(1,2)]
yt_ar2

yt_2_ar <- ar[-c(299,300)]
yt_2_ar
```

```{r}
plot(y=yt_ar2,x=yt_2_ar)
```

```{r}
cor(yt_ar2,yt_2_ar)
```

#### Korelasi antara $Y_t$ dengan $Y_{t-3}$

```{r}
yt_ar3 <- ar[-c(1,2,3)]
yt_ar3

yt_3_ar <- ar[-c(298,299,300)]
yt_3_ar
```

```{r}
plot(y=yt_ar3,x=yt_3_ar)
```

```{r}
cor(yt_ar3,yt_3_ar)
```


# ARMA(2,2)

```{r}
arma.sim <- function(n, macoef, arcoef){
  manum <- length(macoef)
  arnum <- length(arcoef)
  stopifnot(manum < n & arnum < n)
  
  wn <- rnorm(n, sd = 0.5)
  init <- max(manum, arnum)

  arma <- wn[1:init]
  for(i in {init+1}:n){
   mastart <- i - manum
   maend <- i-1
   arstart <- i - arnum
   arend <- i-1
   arma[i] <- sum(arcoef * arma[arstart:arend]) + sum(macoef * wn[mastart:maend])  + wn[i]
   }
  return(arma)
}
```

```{r}
arma22 <- arma.sim(300, c(0.4,0.6), c(0.5,0.2))

arma22 |> arima(c(2,0,2))
```

```{r}
set.seed(1234)
n = length(wn)
phi1 = 0.5
phi2 = 0.2
theta1 = 0.4
theta2 = 0.6

y.arma=c(1:n)
for (i in 3:n){y.arma[i] = phi1*y.arma[i-1] + phi2*y.arma[i-2] + theta1*wn[i-1] + theta2*wn[i-2] + wn[i]}
y.arma
```

```{r}
arma22 <- arima.sim(list(order=c(2,0,2), ar = c(0.5,0.2), ma = c(0.4,0.6)), n=300)
arma22
```

## Karakteristik ARMA(2,2)

### Plot Time Series

```{r}
par(mfrow = c(1, 2))
ts.plot(y.arma)
ts.plot(arma22)
par(mfrow = c(1, 1))
```

### Plot ACF

```{r}
par(mfrow = c(1, 2))
acf(y.arma)
acf(arma22)
par(mfrow = c(1, 1))
```

### Plot PACF

```{r}
par(mfrow = c(1, 2))
pacf(y.arma)
pacf(arma22)
par(mfrow = c(1, 1))
```

### Plot EACF

```{r}
TSA::eacf(y.arma)
TSA::eacf(arma22)
```

### Scatterplot Antar Lag

#### Korelasi antara $Y_t$ dengan $Y_{t-1}$

```{r}
yt_arma <- arma22[-1]
yt_arma

yt_1_arma <- arma22[-300]
yt_1_arma
```

```{r}
plot(y=yt_arma,x=yt_1_arma)
```

```{r}
cor(yt_arma,yt_1_arma)
```

#### Korelasi antara $Y_t$ dengan $Y_{t-2}$

```{r}
yt_arma <- arma22[-c(1,2)]
yt_arma

yt_2_arma <- arma22[-c(299,300)]
yt_2_arma
```

```{r}
plot(y=yt_arma,x=yt_2_arma)
```

```{r}
cor(yt_arma,yt_2_arma)
```

#### Korelasi antara $Y_t$ dengan $Y_{t-3}$

```{r}
yt_arma <- arma22[-c(1,2,3)]
yt_arma

yt_3_arma <- arma22[-c(298,299,300)]
yt_3_arma
```

```{r}
plot(y=yt_arma,x=yt_3_arma)
```

```{r}
cor(yt_arma,yt_3_arma)
```

